Measuring Hash Trustworthiness via Collision Utility Metrics: Logical Cryptanalysis of MD4

### 2017-12-05 dsn-2018/bare_conf.tex

We noted in table \ref{table:neighborhood} that the unit distance neighborhood of
the existing attacks were non-empty. By repeatedly expanding neighborhoods of
known collision classes, we can construct paths in the collision space. We
made the following observations:

\begin{enumerate}
    \item The space of collision classes is not smooth under the unit distance
        relation. That is, starting at $C_{Schlaffer}$ and evaluating
        successive neighborhoods, moving towards $C_{Wang}$ whenever the
        distance decreases does not lead to a simple path of length
        $\delta(C_{Schlaffer}, C_{Wang}) = 17$. Our lower bound on the actual
        distance is 29, but after 14 rounds of expansion, we only reduced the
        distance to 15.
    \item Even under successive unit neighborhoods, the family of input block
        differentials remains roughly the same among successive neighborhoods.
        This can be partially counteracted by moving to reduced-round spaces,
        evaluating the neighborhood, and expanding back to a collision in
        full-round MD4. That is, there is a detectable signature of sharing
        significant input block differential structure with the original
        collision. With 35,918 unique classes of collisions which had 176
        unique families of collisions, there were only three unique families
        of input block differentials: those of the starting collision classes.
\end{enumerate}

This last point shows that, while there may be a number of unique differentials
for a given collision class, and potentially many collision classes within
some neighborhood of the original collision class, by analyzing the structure
of the input block differential, systems can potentially prohibit successful
collision attacks by preventing structures of input block differences.

Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald in
\cite{Schlaffer2006}, this is a 35 fold increase. Increasing this bound is
possible with low additional cost. Further, by fully considering the
similarities of collision families under $\precsim$ and better enumerating the
resulting candidates, additional paths with unique differentials could be
constructed.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{Figs/families48.pdf}
\caption{A visualization of the unit-distance family relationships
of the 35,000 new collisions in MD4. }
\label{Fig:Families48}
\end{center}
\end{figure}

We noted in table \ref{table:neighborhood} that the unit-step neighborhood of
the existing attacks were non-empty. By repeatedly expanding neighborhoods of
known collision classes, we can construct paths in the collision space. We
made the following observations:

The space of collision classes is not smooth under the unit distance
relation. That is, starting at $C_{Schlaffer}$ and evaluating
successive neighborhoods, moving towards $C_{Wang}$ whenever the
distance decreases does not lead to a simple path of length
$\delta(C_{Schlaffer}, C_{Wang}) = 17$. Our lower bound on the actual
distance is 29, but after 14 rounds of expansion, we only reduced the
distance to 15. This shows that $\delta$ does not satisfy the triangle
equality and cannot be a metric, only a heuristic.

Even under successive unit neighborhoods, the family of input block
differentials remains roughly the same among successive neighborhoods,
echoing the work of M. Schl{\"a}ffer and E. Oswald \cite{Schlaffer2006}.
This can be partially counteracted by moving to reduced-round spaces,
evaluating the neighborhood, and expanding back to a collision in
full-round MD4. That is, there is a detectable signature of sharing
significant input block differential structure with the original
collision. Our 35,918 collision classes belonged to 176 unique
families of collisions. However, there were only three unique families
of input block differentials with 100 input block differences. Note
however, that the unlike the metric evaluations in
Section \ref{empirical:differentials}, these collision classes were
not evaluated for number of unique differentials; thus there are likely
more than 100 patterns, but still only three families.

This last point shows that, while there may be a number of unique differentials
for a given collision class, and potentially many collision classes within
some neighborhood of the original collision class, by analyzing the structure
of the input block differential, systems can potentially prohibit successful
collision attacks by preventing structures of input block differences. Note
too how the graph presented in \ref{Fig:Families48} shows a distinct grouping
into three disconnected subgraphs, which fall along input block differential
family boundaries.

Further, compared to the work presented by M. Schl{\"a}ffer and E. Oswald
\cite{Schlaffer2006}, this is a 35 fold increase. Increasing this bound is
possible with low additional cost. Further, by fully considering the
similarities of collision families under $\precsim$ and better enumerating the
resulting candidates, additional paths with unique differentials could be
constructed. This would allow for new candidate input block differentials,
better construction of chosen prefix attacks, and potentially to a full second
preimage attack without relying on message modification techniques.
